Measuring individual differences in the understanding of gaze cues

### 2022-05-23 paper/manuscript.Rmd
incorporate Manuels feedback

  - name          : "Luke Maurits"
    affiliation   : "1"

@@

# Introduction 
<!-- ## What is Social Cognition, and why do we care? -->
Social cognition - representing and reasoning about an agent’s perspectives, knowledge states, intentions, beliefs, and preferences to explain and predict their behavior - is among the most-studied phenomena in developmental research. 
In recent decades, much progress has been made in determining the average age at which a specific social-cognitive ability emerges in development [@gopnik1991young; @wellman2001metaanalysisa; @wellman2004scalinga; @peterson2012mind; @rakoczy2022foundations].

<!-- ## How has it been measured traditionally? -->
One of the most commonly applied prototypical measures for social cognition is the Sally-Anne task [@baron-cohen1985does]. 
Here, children watch a short sequence of events (often acted out or narrated by the experimenters). 
A doll called Sally puts her marble into a basket. 
After Sally leaves the scene, a second doll named Anne takes the marble and moves it into a box. 
Participants then get asked where Sally will look for her marble once she returns. 
The outcome measures false belief understanding in a dichotomous way: 
children pass the task if they take the protagonist’s epistemic state into account and answer that she will look into the basket. 
Many years of research utilizing these verbal change-of-location tasks suggest that children develop belief-representing abilities at four to five years of age [for a review, see @wellman2001metaanalysisa]. 
Several cross-cultural studies supported this evidence [@barrett2013early; @callaghan2005synchrony; cf. @mayer2015weird].
However, from this age onwards, the change-of-location task shows ceiling effects and has very limited diagnostic value. 
As @wellman2012theory put it, “it’s really only passing/failing one sort of understanding averaged across age” (p. 317). 
Therefore, the focus has been to track change at an (age-) group level within a narrow developmental time frame and with a strong emphasis on a single social-cognitive ability.

<!-- ## Why care about individual differences? -->
Although humans substantially differ in their behavior, thoughts, and motivations, researchers often regard variability as cumbersome or reinterpret it as error variance, therefore neglecting and trivializing individual differences [@kidd2018individual]. 
Consequently, little consensus has been reached on the interrelation or independence of social-cognitive abilities [termed ‘structure of social cognition’ by @happe2017structure].

However, identifying variability in social-cognitive abilities and factors influencing those could be vital in theory building (e.g., to test causal predictions) and design interventions [@happe2017structure]. 
For example, those identifiers could be used to provide evidence for two accounts: (1) *developmental lags* (i.e., whether children differ in their speed of acquiring new conceptual insights into the understanding of the mind), and (2) *genuine variation* (i.e., whether children differ in their ease or fluency in (applying) these social-cognitive abilities) [@hughes2015individual]. 
Therefore, we argue that we should study the co-development of social-cognitive abilities, variation in the individual’s pace of mastering developmental milestones, and which factors influence a child’s proficiency. 
For this endeavor, we need to take individual differences seriously – both theoretically and methodologically.

<!-- ## Existing research tapping into individual differences -->
Some studies have already examined individual differences in social cognition [@hughes2015individual; @slaughter2015theorya]. 
Previous studies focused, for example, 
on the association of mental state talk and false belief understanding [@hughes2011individual], 
emotion, language, family interaction, and belief understanding [@dunn1991young; @cutting1999theory; @bulgarelli2016social], 
stereotypes and theory of mind [@rizzo2018theory], 
prosociality and theory of mind [for a review, see @imuta2016theory; @walker2005gender], 
executive functions and theory of mind [@carlson2004individual; @buttelmann2021relations], 
parenting practices and theory of mind [for a review, see @pavarini2013parental], 
agency and moral cognition [@sodian2016understanding], 
attachment quality and gaze following [@astor2020social], 
language and gaze following [@okumura2017individual], 
and the interplay of physical and social cognition [@herrmann2010structure]. 
Correlational approaches have also been used to argue for and test assumptions of cognitive development theories [@mundy2007individual; @kidd2018individual; @underwood1975individual].

<!-- ## Methodological limitations -->
Frequently, developmental psychologists are surprised to find minor or no association between constructs that are thought to be theoretically related [e.g., @sodian2016understanding]. 
This might be because traditional measures of social cognition are not designed to capture variation between children: 
they often rely on low trial numbers, small sample sizes, and dichotomous measures. 
@hedge2018reliability argued that cognitive tasks often become well-established by displaying robust effects across studies. 
The authors could show that this usually does not emerge due to high measurement variance but relatively low between-subject variability, concluding that popular cognitive tasks may reliably measure group differences but not individual differences (termed "reliability paradox"). 
Similarly, @pronk2021methods reasoned that accurate reliability estimates are needed to judge how well a particular cognitive test is suited to draw inferences about individuals.
A recent review underlined reoccurring issues in standardization, validity, reliability, and sensitivity to developmental change: alarmingly few theory of mind measures reported empirically validated psychometric properties [@beaudoin2020systematic] or showed poor test-retest reliability [@mayes1996testretest].

Even studies focusing on individual differences often use behavioral forced-choice paradigms [e.g., carlson2004individual; @walker2005gender]. 
Few studies that do realize continuous measures often administer (theory of mind) scales and conventionally compute aggregate scores (across tasks or trials), therefore neglecting the issue of measurement error [@hughes2011individual]. 
This way, a particular social-cognitive ability in question is still assessed categorically on a trial basis [e.g., @bulgarelli2016social; @rizzo2018theory; @buttelmann2021relations]. 
This has profound implications for what these studies can show. 
Poor measurement of social cognition on an individual level may conceal relations between different aspects of cognition and may obscure developmental change.

<!-- ## A wish list for new social-cognitive tasks -->
Therefore, developmental psychology faces a dilemma: many research questions are questions about individual differences, yet, there is a lack of tasks to measure these individual differences reliably. 
To capture the emergence of social-cognitive abilities and related social factors in greater precision and detail, we must consequently address the methodological limitations of existing study designs [@hughes2004links; @hughes2011individual].

@schaafsma2015deconstructing compiled a "wish-list" for new social-cognitive paradigms. They advocated for parametric measures covering proficiency as a range, avoiding floor and ceiling effects, and showing satisfactory test-retest reliability estimates. 
In valid and reliable tasks, differences in test scores should correspond to differences in the social-cognitive ability and systematically order individuals in the same way. 
Additionally, tasks should allow for use within an fMRI scanner. 
With an easily administrable test procedure, larger sample sizes can be collected to estimate the potentially small effect of inter-individual differences.

Additionally, it has been argued that social-cognitive abilities should not be treated as an all-or-nothing matter (e.g., dichotomously in pass/fail situations) but as abilities on a continuum [@hughes2015individual; @beaudoin2020systematic]. 
Task batteries might be better suited to cover a range of social-cognitive abilities [@hughes2011individual]. 
@schaafsma2015deconstructing proposed that first basic social-cognitive processes should be relatively well understood before advancing to more complex social-cognitive abilities.

Another goal in creating new tasks should be to focus on the face value: 
we want to measure the underlying social-cognitive ability as straight-forward and directly as possible. 
The task should serve as a proxy for behavior as it appears in the real world. 
Importantly, we can assess whether the task shows external validity by evaluating the relationship with related constructs and factors of children's social surroundings.

<!-- ## Our aims -->
Therefore, we aim to make individual differences measurable in a systematic and reliable way. 
We want to create a task that induces variation between children to get precise person-specific estimates. 
To approach this issue, we focus on a continuous measure and short trials that facilitate more than a dozen replicates per subject. 
A standardized, easily accessible test procedure helps us to enable data collection at scale. 
Since data collection with families requires much organizational effort, we want to enable data collection in-person with supervision as well as remotely without supervision.

In our first novel task, we concentrate on the fundamental ability that is involved in many social-cognitive reasoning processes: 
gaze cue understanding - the ability to locate and use the attentional focus of an agent. 
The first component of this ability is often termed gaze following - turning one's eyes in the same direction as the gaze of another agent - and has been studied intensively [@itakura1998use; @delbianco2019developmental; @byers-heinlein2021development; @moore2008development; @coelho2006searching; @tomasello2007reliance; @hernik2019infant; @astor2021gaze; @frischen2007gaze; @shepherd2010following; @lee1998children]. 
Following an agent's gaze provides insights into their intentions, thoughts, and feelings by acting as a 'front end ability' [@brooks2005development, p. 535]. 
In our definition, gaze cue understanding goes one step further by including the *acting on the gaze-cued location* - therefore, using the available social information to guide one's behavior as needed in real-life conditions.

@@

## Implementation
Our balloon finding task is presented as an interactive web app. The task is portable across devices and web browsers and does not require any installation. An advantage of online testing is that our testing procedure is standardized across participants. By using pre-recorded study instructions, no interaction with the experimenter is necessary during the study. The code is open-source (https://github.com/ccp-eva/gafo-demo) and a live demo version can be found under: https://ccp-odc.eva.mpg.de/gafo-demo/.

@@

## Limitations
## Future development / extending the task

>>>>

@@

# Introduction 
Social cognition -- representing and reasoning about an agent’s perspectives, knowledge states, intentions, beliefs, and preferences to explain and predict their behavior -- is among the most-studied phenomena in developmental research. 
In recent decades, much progress has been made in determining the average age at which a specific social-cognitive ability emerges in development [@gopnik1991young; @wellman2001metaanalysis; @wellman2004scaling; @peterson2012mind; @rakoczy2022foundations].
Yet, there are always individual differences. 
Identifying variability in social-cognitive abilities and factors influencing their development is vital in theory building (e.g., to test causal predictions) and for designing interventions 
[@happe2017structure; @mundy2007individual; @kidd2018individual; @underwood1975individual; @lecce2014promoting].

Numerous studies have already examined individual differences in social cognition [for an overview, see @hughes2015individual; @slaughter2015theorya]. 
These individual differences studies often focus on the relationship between social-cognitive abilities and: (1) family influences, (2) other cognitive constructs, and (3) social behavioral outcome [for an overview, see @repacholi2003introduction]. 
Studies on social-cognitive abilities and family influences include the effect of parenting practices [for a review, see @pavarini2013parental], attachment quality [e.g., @astor2020social], mental state talk [@hughes2011individual; @gola2012mental; @lecce2014promoting], and family background as parental education, occupation, sibling interaction and childcare [@dunn1991young; @cutting1999theory; @bulgarelli2016social]. 
Another group of individual differences studies focuses on the interplay of social and physical cognition [@herrmann2010structure], executive functions [@carlson2004individual; @carlson2001individual; @buttelmann2021relations; @hughes2007executive; @benson2012individual], and language abilities [@okumura2017individual; @milligan2007language; @mcewen2007origins]. 
Studies on social behavioral outcomes measured the interplay of social cognition and prosociality [for a review, see @imuta2016theory; @walker2005gender], stereotypes and resource allocations [@rizzo2018theory], and moral intentions [@sodian2016understanding].

However, frequently, developmental psychologists are surprised to find minor or no association between measures of social cognition that are thought to be theoretically related -- cross-sectionally and/or longitudinally [e.g., @sodian2016understanding]. 
This might be because traditional measures of social cognition are not designed to capture variation between children: 
they often rely on low trial numbers, small sample sizes, and dichotomous measures. A recent review showed that many social cognition measures fail to report relevant psychometric properties at all [@beaudoin2020systematic] or -- when they do -- show poor test-retest reliability [@mayes1996testretest].

To give an example: perhaps the most commonly applied prototypical measure for social cognition is the change-of-location false belief task [@baron-cohen1985does; @wimmer1983beliefs]. 
Here, children watch a short sequence of events (often acted out or narrated by the experimenters). 
A doll called Sally puts her marble into a basket. 
After Sally leaves the scene, a second doll named Anne takes the marble and moves it into a box. 
Participants then get asked where Sally will look for her marble once she returns. 
The outcome measures false belief understanding in a dichotomous way: 
children pass the task if they take the protagonist’s epistemic state into account and answer that she will look into the basket. 
Many years of research utilizing these verbal change-of-location tasks suggest that children develop belief-representing abilities at four to five years of age [for a review, see @wellman2001metaanalysis]. 
Several cross-cultural studies supported this evidence [@barrett2013early; @callaghan2005synchrony; cf. @mayer2015weird].

However, from this age onwards, the change-of-location task shows ceiling effects and has very limited diagnostic value [@repacholi2003introduction]. 
Thus, this sort of task seems to be well suited to track group-level developmental trends, yet it fails to capture individual differences [cf. "reliability paradox", @hedge2018reliability]. 
As @wellman2012theory put it, “it’s really only passing/failing one sort of understanding averaged across age” (p. 317). 
This has profound implications for what studies on individual differences using this task (or others) can show. 
Poor measurement of social cognition on an individual level is likely to conceal relations between different aspects of cognition and may obscure developmental change.

Thus, developmental psychology faces a dilemma: many research questions are questions about individual differences, yet, there is a lack of tasks to measure these individual differences reliably. 
To capture the emergence of social-cognitive abilities and related social factors in greater precision and detail, we must consequently address the methodological limitations of existing study designs [@hughes2004links; @hughes2011individual].

@schaafsma2015deconstructing compiled a "wish-list" for new social-cognitive paradigms. 
They advocated for parametric measures covering proficiency as a range, avoiding floor and ceiling effects, and showing satisfactory test-retest reliability estimates [see also @beaudoin2020systematic; @hughes2015individual]. 
New tasks should induce variation across age groups, including older children and adults [@repacholi2003introduction]. 
Another goal in creating new tasks should be to focus on "face value": 
measures should probe the underlying social-cognitive ability as straight-forward and directly as possible. 
The task should serve as a proxy for behavior as it appears in the real world and should be validated in relation to real-world experiences [@repacholi2003introduction].

# A new measure of gaze understanding

Our goal was to design a new measure of social cognition that captures individual differences in a systematic and reliable way. 
We focused on one of the fundamental abilities implicated in many social-cognitive reasoning processes: 
gaze cue understanding -- the ability to locate and use the attentional focus of an agent. 
The first component of this ability is often termed gaze following -- turning one's eyes in the same direction as the gaze of another agent -- 
and has been studied intensively [@itakura1998use; @delbianco2019developmental; @byers-heinlein2021development; @moore2008development; @coelho2006searching; @tomasello2007reliance; @hernik2019infant; @astor2021gaze; @frischen2007gaze; @shepherd2010following; @lee1998children]. 
Following an agent's gaze provides insights into their intentions, thoughts, and feelings by acting as a "front end ability" [@brooks2005development, p. 535]. 
In our definition, gaze cue understanding goes one step further by including the *acting on the gaze-cued location* -- therefore, using the available social information to guide one's behavior as needed in real-life conditions.

To address the psychometric shortcoming of earlier work, we implemented the following design features: 
First, we used a continuous measure which allowed us to capture fine grained individual differences at different ages. 
Second, we designed short trials that facilitate more than a dozen replicates per subject. 
The result are more precise individual-level estimates. 
Third, we systematically investigated the psychometric properties of the new task. 
Finally, to validate the task, we studied how it relates to aspects of children's everyday experience.

Designing this task required a new testing infrastructure. 
We designed the task as an interactive browser-based web app. 
This greatly increased the flexibility with which we could modify the stimuli on a trial-by-trial basis. 
Furthermore, because the task is largely self-contained, it is much more controlled and standardized. 
Most importantly perhaps, it makes the task portable: testing is possible in-person using tablets but also remotely via the internet (no installation needed). 
As such, it provides a solid basis to study individual differences in gaze cue understanding across ages at scale. 
We make the task and its source code openly accessible for other researchers to use and modify.

@@

## Implementation
The code is open-source (https://github.com/ccp-eva/gafo-demo) and a live demo version can be found under: https://ccp-odc.eva.mpg.de/gafo-demo/.

@@

- benefits of hedge (psychometric properties) vs box (discrete, easier for kids to click, induce salience etc)

From previous intro potentially: 
- @hedge2018reliability section on reliability paradox
- Few studies that do realize continuous measures often administer (theory of mind) scales and conventionally compute aggregate scores (across tasks or trials), therefore neglecting the issue of measurement error [@hughes2011individual]. 
This way, a particular social-cognitive ability in question is still assessed categorically on a trial basis [e.g., @bulgarelli2016social; @rizzo2018theory; @buttelmann2021relations]. 
- Additionally, it has been argued that social-cognitive abilities should not be treated as an all-or-nothing matter (e.g., dichotomously in pass/fail situations) but as abilities on a continuum [@hughes2015individual; @beaudoin2020systematic]. 
Task batteries might be better suited to cover a range of social-cognitive abilities [@hughes2011individual].

## Limitations
## Future development / extending the task
- benefits of task battery
